03-modeling

### 2018-04-26 02_scripts/01_hyperspectral_indices/06_extract-indices-to-points.Rmd
26/04/18

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE)
```

```{r, results='hide'}
library(raster)
devtools::install_github("pat-s/hsdar")
library(hsdar)
library(pbmcapply)
library(glue)
library(magrittr)
library(sf)
library(purrr)
```

# Laukiz1

## Load points

```{r}
st_read("/home/shares/data/LIFE/mod/tree-per-tree/laukiz1/laukiz1.shp", 
        quiet = TRUE) %>% 
  as("Spatial") -> laukiz1_points
```

## Load rasters

```{r}
laukiz1_veg_indices <- brick("/home/shares/data/LIFE/mod/hyperspectral/all-veg-indices/all-veg-indices-laukiz1")
```

```{r}
laukiz1_nbi <- brick("/home/shares/data/LIFE/mod/hyperspectral/narrow-band-indices/narrow-band-indices-laukiz1")
```

## Extract values to points

### Veg indices

Buffer size:

* 1m
* 2m
* 3m
* 4m
* 5m

```{r}
out_veg <- lapply(seq_along(1:5), function(x) 
  raster::extract(laukiz1_veg_indices, laukiz1_points, buffer = x, 
                  fun = mean, df = TRUE))
```

Label output names according to buffer size

```{r}
out_veg <- lapply(seq_along(out_veg), function(x) 
  setNames(out_veg[[x]], paste0("bf", x, "_", colnames(out_veg[[x]]))))
```

### NBI

This might take a while - nested parallelization here! 
Only outer parallelization: 1 hour

This can only be run on a server infrastructure! This code uses 8*5 = 40 + 5 = 45 cores!

NOTE 05 April: Nested parallelization causes some weird errors for Laukiz2 and does not really speed up the processing. 
I think we are better of using only outer parallelization here.

```{r}
if (detectCores() > 45 ) {

  Sys.time <- Sys.time()

  out_nbi <- pblapply(seq_along(1:5), function(x) {
    #beginCluster(8) # assigning 8 cores to each parallelized `extract` call -> 40 cores total because 8*5
    raster::extract(laukiz1_nbi, laukiz1_points, buffer = x, 
                    fun = mean, df = TRUE) } , cl = 5)

  print(Sys.time() - Sys.time)
} else {
  stop("please run this script on a server with more than 45 cores!")
}
```

### Bind to points

```{r}
# merge all data frames
df_out <- bind_cols(out_veg)

df_out1 <- bind_cols(out_nbi)

laukiz1_points %>% 
  st_as_sf() %>% 
  bind_cols(df_out) %>% 
  bind_cols(df_out1) -> laukiz1_points
```

## Save files

```{r}
# .rda file (data frame)
as.data.frame(laukiz1_points) %>% 
  saveRDS("/home/shares/data/LIFE/mod/tree-per-tree/laukiz1/all-indices-laukiz1.rda")
```

# Laukiz2

## Load points

```{r}
print("Starting Laukiz2")
```

```{r}
st_read("/home/shares/data/LIFE/mod/tree-per-tree/laukiz2/laukiz2.shp", 
        quiet = TRUE) %>% 
  as("Spatial") -> laukiz2_points
```

## Load rasters

```{r}
laukiz2_veg_indices <- brick("/home/shares/data/LIFE/mod/hyperspectral/all-veg-indices/all-veg-indices-laukiz2")
```

```{r}
laukiz2_nbi <- brick("/home/shares/data/LIFE/mod/hyperspectral/narrow-band-indices/narrow-band-indices-laukiz2")
```

## Extract values to points

### Veg indices

Buffer size:

* 1m
* 2m
* 3m
* 4m
* 5m

```{r}
out_veg <- lapply(seq_along(1:5), function(x) 
  raster::extract(laukiz2_veg_indices, laukiz2_points, buffer = x, 
                  fun = mean, df = TRUE))
```

Label output names according to buffer size

```{r}
out_veg <- lapply(seq_along(out_veg), function(x) 
  setNames(out_veg[[x]], paste0("bf", x, "_", colnames(out_veg[[x]]))))
```

### NBI

This might take a while - nested parallelization here! 
Only outer parallelization: 1 hour

This can only be run on a server infrastructure! This code uses 8*5 = 40 + 5 = 45 cores!

NOTE 05 April: Neste parallelization causes some weird errors for Laukiz2 and does not really speed up the processing. 
I think we are better of using only outer parallelization here.

```{r}
if (detectCores() > 45 ) {

  Sys.time <- Sys.time()

  out_nbi <- pblapply(seq_along(1:5), function(x) {
    #beginCluster(8) # assigning 8 cores to each parallelized `extract` call -> 40 cores total because 8*5
    raster::extract(laukiz2_nbi, laukiz2_points, buffer = x, 
                    fun = mean, df = TRUE) } , cl = 5)
  #endCluster()
  print(Sys.time() - Sys.time)
} else {
  stop("please run this script on a server with more than 45 cores!")
}

### local debugging code
# out_nbi <- raster::extract(laukiz2_nbi, laukiz2_points, buffer = 2, 
#                            fun = mean, df = TRUE)
```

### Bind to points

```{r}
# merge all data frames
df_out <- bind_cols(out_veg)

df_out1 <- bind_cols(out_nbi)

laukiz2_points %>% 
  st_as_sf() %>% 
  bind_cols(df_out) %>% 
  bind_cols(df_out1) -> laukiz2_points
```

## Save files

```{r}
# .rda file (data frame)
as.data.frame(laukiz2_points) %>% 
  saveRDS("/home/shares/data/LIFE/mod/tree-per-tree/laukiz2/all-indices-laukiz2.rda")
```

# Oiartzun

## Load points

```{r}
print("Starting oiartzun")
```

```{r}
st_read("/home/shares/data/LIFE/mod/tree-per-tree/oiartzun/oiartzun.shp", 
        quiet = TRUE) %>% 
  as("Spatial") -> oiartzun_points
```

## Load rasters

```{r}
oiartzun_veg_indices <- brick("/home/shares/data/LIFE/mod/hyperspectral/all-veg-indices/all-veg-indices-oiartzun")
```

```{r}
oiartzun_nbi <- brick("/home/shares/data/LIFE/mod/hyperspectral/narrow-band-indices/narrow-band-indices-oiartzun")
```

## Extract values to points

### Veg indices

Buffer size:

* 1m
* 2m
* 3m
* 4m
* 5m

```{r}
out_veg <- lapply(seq_along(1:5), function(x) 
  raster::extract(oiartzun_veg_indices, oiartzun_points, buffer = x, 
                  fun = mean, df = TRUE))
```

Label output names according to buffer size

```{r}
out_veg <- lapply(seq_along(out_veg), function(x) 
  setNames(out_veg[[x]], paste0("bf", x, "_", colnames(out_veg[[x]]))))
```

### NBI

This might take a while - nested parallelization here! 
Only outer parallelization: 1 hour

This can only be run on a server infrastructure! This code uses 8*5 = 40 + 5 = 45 cores!

NOTE 05 April: Neste parallelization causes some weird errors for Laukiz2 and does not really speed up the processing. 
I think we are better of using only outer parallelization here.

```{r}
if (detectCores() > 45 ) {

  Sys.time <- Sys.time()

  out_nbi <- pblapply(seq_along(1:5), function(x) {
    #beginCluster(8) # assigning 8 cores to each parallelized `extract` call -> 40 cores total because 8*5
    raster::extract(oiartzun_nbi, oiartzun_points, buffer = x, 
                    fun = mean, df = TRUE) } , cl = cores)

  print(Sys.time() - Sys.time)
} else {
  stop("please run this script on a server with more than 45 cores!")
}
```

### Bind to points

```{r}
# merge all data frames
df_out <- bind_cols(out_veg)

df_out1 <- bind_cols(out_nbi)

oiartzun_points %>% 
  st_as_sf() %>% 
  bind_cols(df_out) %>% 
  bind_cols(df_out1) -> oiartzun_points
```

## Save files

```{r}
# .rda file (data frame)
as.data.frame(oiartzun_points) %>% 
  saveRDS("/home/shares/data/LIFE/mod/tree-per-tree/oiartzun/all-indices-oiartzun.rda")
```

# Luiando

## Load points

```{r}
print("Starting luiando")
```

```{r}
st_read("/home/shares/data/LIFE/mod/tree-per-tree/luiando/luiando.shp", 
        quiet = TRUE) %>% 
  as("Spatial") -> luiando_points
```

## Load rasters

```{r}
luiando_veg_indices <- brick("/home/shares/data/LIFE/mod/hyperspectral/all-veg-indices/all-veg-indices-ayala-aiara") 
```

```{r}
luiando_nbi <- brick("/home/shares/data/LIFE/mod/hyperspectral/narrow-band-indices/narrow-band-indices-ayala-aiara")
```

## Extract values to points

### Veg indices

Buffer size:

* 1m
* 2m
* 3m
* 4m
* 5m

```{r}
out_veg <- lapply(seq_along(1:5), function(x) 
  raster::extract(luiando_veg_indices, luiando_points, buffer = x, 
                  fun = mean, df = TRUE))
```

Label output names according to buffer size

```{r}
out_veg <- lapply(seq_along(out_veg), function(x) 
  setNames(out_veg[[x]], paste0("bf", x, "_", colnames(out_veg[[x]]))))
```

### NBI

This might take a while - nested parallelization here! 
Only outer parallelization: 1 hour

This can only be run on a server infrastructure! This code uses 8*5 = 40 + 5 = 45 cores!

NOTE 05 April: Neste parallelization causes some weird errors for Laukiz2 and does not really speed up the processing. 
I think we are better of using only outer parallelization here.

```{r}
if (detectCores() > 45 ) {

  Sys.time <- Sys.time()

  out_nbi <- pblapply(seq_along(1:5), function(x) {
    #beginCluster(8) # assigning 8 cores to each parallelized `extract` call -> 40 cores total because 8*5
    raster::extract(luiando_nbi, luiando_points, buffer = x, 
                    fun = mean, df = TRUE) } , cl = cores)

  print(Sys.time() - Sys.time)
} else {
  stop("please run this script on a server with more than 45 cores!")
}
```

### Bind to points

```{r}
# merge all data frames
df_out <- bind_cols(out_veg)

df_out1 <- bind_cols(out_nbi)

luiando_points %>% 
  st_as_sf() %>% 
  bind_cols(df_out) %>% 
  bind_cols(df_out1) -> luiando_points
```

## Save files

```{r}
# .rda file (data frame)
as.data.frame(luiando_points) %>% 
  saveRDS("/home/shares/data/LIFE/mod/tree-per-tree/luiando/all-indices-luiando.rda")
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r, results='hide'}
library(raster)
devtools::install_github("pat-s/hsdar")
library(hsdar)
library(pbmcapply)
library(glue)
library(magrittr)
library(sf)
library(purrr)
library(dplyr)
```

The points to which the indices are extracted to are cleaned in the file "02_preprocessing.Rmd".

# Create function

See step-by-step example for Laukiz 1 below.

```{r}
extract_to_plot = function(plot_name, buffer, cores, bf_name) {
  points = st_read(glue("/data/patrick/mod/tree-per-tree/2016/{plot_name}/{plot_name}.gpkg"),
          quiet = TRUE)

  ras_veg_indices <- brick(glue("/data/patrick/mod/hyperspectral/all-veg-indices/{plot_name}.grd"))

  ras_nbi <- brick(glue("/data/patrick/mod/hyperspectral/narrow-band-indices/narrow-band-indices-{plot_name}.grd"))

  # calculate buffered veg index
  out_veg <- pbmclapply(buffer, function(x) 
    raster::extract(ras_veg_indices, points, buffer = x, 
                    fun = mean, df = TRUE,
                    na.rm = TRUE), mc.cores = cores)

  out_veg %<>% 
    map2(seq_along(buffer), ~ setNames(.x, glue("bf", .y, "_{name}", 
                                             name = names(out_veg[[.y]]))))

  out_nbi <- pbmclapply(buffer, function(x)
    raster::extract(ras_nbi, points, buffer = x,
                    fun = mean, df = TRUE,
                    na.rm = TRUE) , mc.cores = cores)

  out_nbi %<>%
    map2(seq_along(buffer), ~ setNames(.x, glue("bf", .y, "_{name}",
                                             name = names(out_nbi[[.y]]))))

  # merge all data frames (buffers)
  all_veg <- bind_cols(out_veg)

  all_nbi <- bind_cols(out_nbi)

  points %<>% 
    bind_cols(all_veg) %>% 
    bind_cols(all_nbi)

  # .rda file (data frame)
  points %>% 
    saveRDS(glue("/data/patrick/mod/tree-per-tree/2016/{plot_name}/all-indices-{plot_name}_{bf_name}.rda"))
}
```

# Process

## Buffer: 2

```{r}
pbmclapply(c("laukiz1", "laukiz2", "oiartzun", "luiando"), 
           function(x) extract_to_plot(x, 2, cores = 1, "bf2"), 
           mc.cores = 4)
```

<!-- ## Buffer: 1 - 10 -->

<!-- ```{r} -->
<!-- pbmclapply(c("laukiz1", "laukiz2", "oiartzun", "luiando"),  -->
<!--            function(x) extract_to_plot(x, seq(1:10), cores = 10, "bf1-10"),  -->
<!--            mc.cores = 4) -->
<!-- ``` -->

<!-- ## Buffer: 1,5,10,15,20 -->

<!-- ```{r} -->
<!-- pbmclapply(c("laukiz1", "laukiz2", "oiartzun", "luiando"),  -->
<!--            function(x) extract_to_plot(x, c(1, 5, 10, 15, 20), cores = 5, "bf1-5-10-15-20"),  -->
<!--            mc.cores = 4) -->
<!-- ``` -->

<!-- ## Buffer: 1, 10, 20, 30, 40, 50 -->

<!-- ```{r} -->
<!-- pbmclapply(c("laukiz1", "laukiz2", "oiartzun", "luiando"),  -->
<!--            function(x) extract_to_plot(x, c(1, 10, 20, 30, 40, 50), cores = 6,  -->
<!--                                        "bf1-10-20-30-40-50"),  -->
<!--            mc.cores = 2) # due to RAM consumption -->
<!-- ``` -->

### 2018-04-26 02_scripts/032-laukiz2/03-modeling_laukiz2.Rmd
26/04/18

```{r}
data = data[[1]]
coords = coords[[1]]
```

@@

```{r}
# predict.type for 'auc' (AUROC)
lrn_glmnet <- makeLearner("regr.glmnet",
                          alpha = 0,
                          standardize = FALSE,
                          intercept = FALSE)
                          #lambda = sort(seq(0, 5, length.out = 100), decreasing = T))
# getHyperPars(lrn_glmnet)
# getParamSet(lrn_glmnet)
```

@@

```{r}
ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 77, upper = 7700))
```

```{r}
ctrl <- makeTuneControlRandom(maxit = 200)
inner <- makeResampleDesc("SpCV", iters = 10)
```

@@

```{r}
ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.07, upper = 7.7))
```

@@

```{r}
outer <- makeResampleDesc("SpRepCV", folds = 5, reps = 2)
```

```{r}
library(parallelMap)
parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
              mc.set.seed = TRUE) # only parallelize the tuning

@@

For the elasticnet we also tune `alpha`.
As we have both penalties, we just use the lower and uppper maxima from both Lasso and Ridge.

```{r}
ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.07, upper = 7700),
                          makeNumericParam("alpha", lower = 0, upper = 1))
```

```{r}
ctrl <- makeTuneControlRandom(maxit = 200)
inner <- makeResampleDesc("SpCV", iters = 10)
```

@@

```{r}
configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
library(parallelMap)
parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
              mc.set.seed = TRUE) # only parallelize the tuning
set.seed(12345)
params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
                                 par.set = ps_glmnet, control = ctrl, 
                                 measure = list(rmse))
parallelStop()
```

>>>>

```{r}
data = data[[2]]
coords = coords[[2]]
```

@@

```{r}
# predict.type for 'auc' (AUROC)
lrn_glmnet <- makeLearner("regr.glmnet",
                          alpha = 0,
                          standardize = FALSE)
                          #lambda = sort(seq(0, 5, length.out = 100), decreasing = T))
getHyperPars(lrn_glmnet)
getParamSet(lrn_glmnet)
```

@@

```{r}
ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 30, upper = 3300))
```

```{r}
ctrl <- makeTuneControlRandom(maxit = 500)
inner <- makeResampleDesc("SpCV", iters = 5)
```

@@

```{r}
ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.03, upper = 3.3))
```

```{r}
ctrl <- makeTuneControlRandom(maxit = 500)
inner <- makeResampleDesc("SpCV", iters = 5)
```

@@

@@

For the elasticnet we also tune `alpha`.

```{r}
ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.03, upper = 3.3),
                          makeNumericParam("alpha", lower = 0, upper = 1))
```

@@

### 2018-04-26 02_scripts/035_all-plots/03-modeling_all.Rmd
26/04/18

```{r}
data = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2.rda")
coords = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2-coords.rda")
```

# Select plot

```{r}
data = data[[1]]
coords = coords[[1]]
```

@@

We only need to tune `s`: https://github.com/mlr-org/mlr/issues/1030#issuecomment-233677172
`glmnet` fits many models during training. However, important is which `s` is chosen for prediction. 
The value of `s` determines which trained model will be chosen.
If `s` != `lambda`, linear interpolation or refitting (this) is applied.

@@

Coefficients of ridge regression should not be used for inference as they are biased. 
They can be seen as a variable importance measure. 
The highest coefficients are most important for the model (if all variables are on the same scale) https://stats.stackexchange.com/a/154712/101464

@@

set.seed(12345)
resa_glmnet <- mlr::resample(tune_wrapper, task,
                             resampling = outer, extract = getTuneResult,
                             show.info = TRUE, measures = list(rmse))
parallelStop()
saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_ridge.rda")
```

@@

```{r}
lrn_glmnet_tuned <- makeLearner("regr.glmnet",
                                alpha = 0,
                                s = 1810,
                                standardize = FALSE, 
                                intercept = FALSE)

glmnet_train_tuned = mlr::train(lrn_glmnet_tuned, task)
```

@@

@@

```{r}
lrn_glmnet <- makeLearner("regr.glmnet",
                          alpha = 1,
                          standardize = FALSE,
                          intercept = FALSE)
# getHyperPars(lrn_glmnet)
# getParamSet(lrn_glmnet)
```

@@

parallelStop()
saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_lasso.rda")
```

```{r}
plot(glmnet_train_tuned$learner.model, "norm")
```

### Elasticnet

```{r}
lrn_glmnet <- makeLearner("regr.glmnet",
                          standardize = FALSE,
                          intercept = FALSE)
#getHyperPars(lrn_glmnet)
#getParamSet(lrn_glmnet)
```

For the elasticnet we also tune `alpha`.

```{r}
ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.03, upper = 3.3),
                          makeNumericParam("alpha", lower = 0, upper = 1))
```

```{r}
ctrl <- makeTuneControlRandom(maxit = 200)
inner <- makeResampleDesc("SpCV", iters = 5)
```

parallelStop()
saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_elasticnet")
```

>>>>

```{r}
data = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf1.rda")
coords = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf1-coords.rda")
```

# Merge into one dataset

Leaving out Luiando (see README for reason).

```{r}
data = as_tibble(rbindlist(data, fill = TRUE))
data = Filter(function(x) !any(is.na(x)), data)
coords = as_tibble(rbindlist(coords))
```

@@

@@

@@

parallelStop()
```

@@

@@

#### Variable importance

```{r}
library(parallelMap)
parallelStart(mode = "multicore", cpus = 40,
              mc.set.seed = TRUE) # only parallelize the tuning
set.seed(12345)
feature_imp = generateFeatureImportanceData(task, 
                                            method = "permutation.importance",
                                            lrn_glmnet_tuned,
                                            measure = list(rmse),
                                            nmc = 40)
parallelStop()

saveRDS(feature_imp, "/data/patrick/mod/hyperspectral/variable-importance/bf1-nmc40.rda")

arrange(feature_imp$res)

melt(feature_imp$res) %>% 
  arrange(desc(value))
```

@@

```{r}
# predict.type for 'auc' (AUROC)
lrn_glmnet <- makeLearner("regr.glmnet",
                          alpha = 1,
                          standardize = FALSE)
                          #lambda = sort(seq(0, 5, length.out = 100), decreasing = T))
getHyperPars(lrn_glmnet)
getParamSet(lrn_glmnet)
```

@@

parallelStop()
saveRDS("/data/patrick/mod/tmp/tuning_hyperspectral/tuning_bf1_lasso.rda")
```
